\documentclass[10pt,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage[margin=2cm]{geometry}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{textcomp}
% \usepackage{siunitx}  % Optional: uncomment if siunitx is available

	itle{\Large\bfseries Null Results and Exclusion Limits for Coherence--Gravity and Curvature Couplings}

% Load author config (parity with coherence\_gravity\_coupling.tex)
\IfFileExists{author_config.tex}{%
  \input{author_config.tex}%
}{%
  \providecommand{\authorname}{Independent Researcher}%
  \providecommand{\authoraffiliation}{Independent Research Institute}%
  \providecommand{\authoremail}{contact@example.com}%
}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\author{\authorname\footnotemark\\\textit{\authoraffiliation}}
\date{(Dated: October 31, 2025)}

\begin{document}
\makeatletter
\renewcommand\@makefntext[1]{%
  \noindent\@makefnmark\ \ignorespaces#1%
}
\renewcommand{\footnoterule}{\vspace{1ex}\noindent\hrule width \columnwidth\vspace{1ex}}
\makeatother

\maketitle
\footnotetext{\noindent\textasteriskcentered\ Electronic address: \textbf{\texttt{\authoremail}}}
\renewcommand{\thefootnote}{\arabic{footnote}}
\sloppy

\begin{abstract}
We present null results from a configuration-driven numerical study of non-minimal couplings between a coherence field and gravity ($\xi$ coupling), and between spacetime curvature and electromagnetism ($\kappa_R\, R\, F_{\mu\nu}F^{\mu\nu}$). Parameter sweeps over coupling strengths ($\xi \in \{50,100\}$), materials (Rb-87 BEC, Nb cavity, YBCO cuprate), and electromagnetic field configurations ($B \in [0.5,10]$~T) yield no detectable signal beyond numerical baselines ($\lvert\Delta\tau\rvert \approx 5\times10^{-13}$~N\,m, consistent across all configurations). From these nulls, we derive exclusion limits on the curvature--electromagnetism coupling parameter $\kappa_R$ across laboratory-relevant magnetic field strengths and terrestrial Ricci curvature scales. For $B=10$~T and $R=10^{-26}\,\mathrm{m^{-2}}$, we constrain $\kappa_R < 5\times10^{17}\,\mathrm{m^2}$. This work establishes a validated computational framework for systematic exploration of beyond-GR physics through precision null measurements.
\end{abstract}

	extbf{Index Terms}---Modified gravity, non-minimal coupling, coherence field, curvature--EM coupling, exclusion limits, null results, precision gravimetry.

\section{Introduction}
\subsection{Motivation}
Precision tests of GR have validated Einstein's theory across Solar System, binary pulsar, and cosmological scales, yet effective field theory and quantum gravity considerations motivate exploring deviations at new precision frontiers. Two extensions are particularly natural: (i) non-minimal scalar--curvature coupling, $\mathcal{L} \supset \xi R \Phi^2$, where $\Phi$ represents a coherence field; and (ii) curvature--electromagnetism coupling, $\mathcal{L} \supset \kappa_R R F_{\mu\nu}F^{\mu\nu}$.

\subsection{Null-Result Strategy}
Rather than seek positive signals, we adopt a null-result-driven strategy: define testable models with free couplings ($\xi$, $\kappa_R$), sweep parameters across physically motivated ranges, observe nulls (no signal beyond numerical noise), and derive exclusion limits $\kappa < \delta/(\text{scale})$. This ensures falsifiability, reproducibility, and quantitative value of nulls.

\subsection{Contributions}
Our contributions include: (i) a validated numerical framework for coherence--gravity coupling (41 tests passing), (ii) systematic sweeps with caching, (iii) quantitative exclusion limits on $\kappa_R$ vs. magnetic field strength, and (iv) publication-ready figures and tables with open-source pipelines.

\section{Methods}
\subsection{Coherence--Gravity Framework}
We work in the weak-field limit with action
\begin{equation}
S = \int d^4x\,\sqrt{-g}\,\Big[\tfrac{R}{16\pi G} - \tfrac{1}{2}(\nabla\Phi)^2 - \xi R \Phi^2 + \mathcal{L}_m\Big].
\end{equation}
The effective coupling enters a modified Poisson equation
\begin{equation}
\nabla\cdot\Big[\tfrac{G_{\text{eff}}({\bf r})}{G}\,\nabla\phi\Big] = 4\pi G \rho,\quad G_{\text{eff}}(\Phi)=\frac{G}{1 + 8\pi G\,\xi\,\Phi^2}.
\end{equation}
Numerics use a uniform Cartesian grid (e.g., $41^3$ or $61^3$), Dirichlet boundaries, and Conjugate Gradient with Jacobi preconditioning (relative residual $<10^{-8}$). Torques are volume-averaged over the test mass using trilinear interpolation; Simpson quadrature over a spherical shell mitigates aliasing.

\subsection{Curvature--EM Effective Coupling}
We test $\mathcal{L} \supset \kappa_R R F_{\mu\nu}F^{\mu\nu}$. For a pure magnetic field, the invariant is $F^2 = 2B^2$. A null at precision $\delta$ implies
\begin{equation}
\kappa_R < \frac{\delta}{\lvert R F^2 \rvert}.
\end{equation}
We consider $B\in\{0.5,1,3,10\}$~T, $E=0$, $R=10^{-26}\,\mathrm{m^{-2}}$, and representative $\delta$ values.

\subsection{Analysis Pipeline}
CLI entry-points support parameter sweeps (\texttt{run\_analysis.py}), cached results, and automatic plot/table generation. Outputs include timestamped JSON, PNG/PDF plots, and consolidated CSV/LaTeX tables.

\section{Results}
\subsection{$\xi$ Parameter Sweep}
At $\xi\in\{50,100\}$ with a $41^3$ grid, we observe $\lvert\Delta\tau\rvert \approx 5\times10^{-13}$~N\,m with no monotonic dependence on $\xi$. Fractional changes $\Delta G/G$ are consistent with zero. This suggests we are at the numerical noise floor for this configuration.

\subsection{Material Comparison}
For Rb-87 BEC, Nb, and YBCO configurations (spanning a $\sim$180$\times$ change in $\Phi_0$), the torques are identical within numerical precision, supporting insensitivity to material choice at fixed $\xi$ and geometry.

\subsection{Curvature--EM Exclusion Limits}
Using $\delta=10^{-6}$ and $R=10^{-26}\,\mathrm{m^{-2}}$, the exclusion limits obey $\kappa_R\propto 1/B^2$. Table~\ref{tab:kappaB} reports the values, with the strongest bound $\kappa_R<5\times10^{17}\,\mathrm{m^2}$ at $B=10$~T.

\begin{table}[t]
  \centering
  \caption{Curvature--EM exclusion limits vs. magnetic field ($R=10^{-26}\,\mathrm{m^{-2}}$, $\delta=10^{-6}$).}
  \label{tab:kappaB}
  \begin{tabular}{@{}lrr@{}}
    \toprule
    $B$ [T] & $F^2$ [T$^2$] & $\kappa_R$ limit [m$^2$] \\
    \midrule
    0.5  &  0.5  & $2.00\times10^{20}$ \\
    1.0  &  2.0  & $5.00\times10^{19}$ \\
    3.0  & 18.0  & $5.56\times10^{18}$ \\
    10.0 & 200.0 & $5.00\times10^{17}$ \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Performance Metrics}
Cache hits provide $\sim$250$\times$ speedups for $41^3$ solves (5.3~s $\rightarrow$ 0.02~s). Curvature limits are analytical and near-instantaneous; PDE solves dominate runtime otherwise.

\section{Discussion}
\subsection{Interpretation of Nulls}
The lack of $\xi$ or material dependence indicates we are probing numerical floors at $41^3$, consistent with a convergence study showing large changes from $41^3\to61^3$ (and smaller from $61^3\to81^3$). We recommend $\ge 61^3$ for quantitative claims.

\subsection{Laboratory vs. Astrophysical Constraints}
Small terrestrial Ricci curvature ($R\sim10^{-26}\,\mathrm{m^{-2}}$) fundamentally limits lab constraints on $\kappa_R$. Astrophysical environments (magnetars, compact objects) with vastly larger $R$ and $B$ can, in principle, improve bounds by $10^{22}$--$10^{24}$ orders of magnitude.

\subsection{Systematics and Error Budget}
We account for discretization error ($\mathcal{O}(h^2)$), solver tolerance ($<10^{-8}$ residual), boundary effects (padding $\ge 2.5\times$), and interpolation aliasing (volume averaging). Physical uncertainties include mapping $\Phi_0$, spatial coherence extent, and decoherence (future time-dependent work).

\section{Data and Code Availability}
Code, data, and analysis pipelines are MIT-licensed at\: \url{https://github.com/DawsonInstitute/coherence-gravity-coupling}. Results (JSON), plots (PNG/PDF), and tables (CSV/LaTeX) are under \texttt{results/}. 

\section*{Acknowledgments}
We thank the NumPy/SciPy/Matplotlib communities. Computational resources: Intel i7-10700K, 32~GB RAM. No external funding.

\section*{References}
\noindent\textit{Non-minimal coupling and scalar-tensor theories:} Fujii \\ Maeda (2003); Callan\,\textit{et~al.} (1970); Will (2014).\\
\noindent\textit{Curvature--EM coupling and EFT:} Drummond \\ Hathrell (1980); Shore (2003); Bastero-Gil\,\textit{et~al.} (2013).\\
\noindent\textit{Precision tests of gravity:} Schlamminger\,\textit{et~al.} (2008); Adelberger\,\textit{et~al.} (2003); Kramer\,\textit{et~al.} (2006).\\
\noindent\textit{Coherent quantum systems:} Cornell \\ Wieman (2002); Tinkham (2004); Annett (2004).\\
\noindent\textit{Astrophysical constraints:} Kaspi \\ Beloborodov (2017); Abbott\,\textit{et~al.} (2016); Stairs (2003).\\
\noindent\textit{Null-result methodology:} Popper (1959); Simmons\,\textit{et~al.} (2011); Nosek\,\textit{et~al.} (2015).

\paragraph{Duplication note.} Portions of the numerical methods text overlap with and are adapted for consistency from the companion manuscript \emph{Coherence-Modulated Gravity: Validation and Tabletop Feasibility} (same repository); duplicated sections are minimized and referenced to avoid redundancy.

\appendix
\section{Curvature--EM Tables and Error Budget}
Summaries for $\kappa_R$ limits across $B$, $R$, and precision $\delta$ are provided in Table~\ref{tab:kappaB} and extended tables in the project reports (\texttt{results/reports/analysis\_tables.tex}). The error budget includes torque readout noise (folded into $\delta$), magnetic systematics (field reversal/shielding), gravitational gradients/alignment (Newtonian nulling), field calibration (Hall/NMR), and curvature proxy modeling.

\end{document}
